### 2016-08-22 robust-software.tex
Filling in prior to asking for feedback.
1. Filled in section on making old versions available.
2. Moved section on dependencies (since it seems to fit most naturally right after the old version section).
3. Rewrote conclusion to emphasize changing the ratio between effort available and effort required.

\begin{flushleft}
{\Large
\textbf{Ten Simple Rules for Making Research Software More Robust}
}

{Morgan~Taschuk}$^{1,\ast}$,
{Greg~Wilson}$^{2}$
\\
1) Ontario Institute for Cancer Research / morgan.taschuk@oicr.on.ca
\\
2) Software Carpentry Foundation / gvwilson@software-carpentry.org
\\
$\ast$ Corresponding author.
\end{flushleft}

\begin{quote}
\begin{center}\textbf{Abstract}\end{center}
Software produced for research---even published software---suffers from a number of common problems
that make it difficult or impossible to run outside the original institution
or even the primary developer's computer.
We present ten simple rules to make software robust enough to run anywhere
and delight your users and collaborators.
\end{quote}

@@

Most software has a version number composed of a decimal number that
increments as new versions are released. There are many different ways
to construct and interpret the version, but most importantly for us, a
particular software version run with the same parameters should give
identical results no matter when it's run. Results include both correct
output as well as any errors.
Every time you release your software, i.e.~distribute it to
someone other than yourself and/or the development team, you should
increment your version number.

\withurl{Semantic versioning}{http://semver.org/} is one of the most common
types of versioning for open-source software. Version numbers take the
form of \emph{MAJOR.MINOR{[}.PATCH{]}}, e.g.~0.2.6. The major and
minor numbers versions are almost always provided. Changes in the major
version number herald significant changes in the software that are not
backwards compatible, such as changing or removing features or altering
the primary functions of the software. Increasing the minor version
represents incremental improvements in the software, like adding new
features. Following the minor version number can be an arbitrary number
of project-specific identifiers, including patches, builds and qualifiers.
Common qualifiers include \texttt{alpha}, \texttt{beta}, and \texttt{SNAPSHOT},
for applications that are
not yet stable or released, and \texttt{-RC} for release candidates prior
to official release.

@@

\fixme{MT}{Can you please flesh this out?}

@@

Useful parameters are those values that a user will need to modify to suit
their computer, dataset or application. 
The full list of useful parameters is software-specific and so cannot be
detailed here, but here is a short list:

\begin{itemize}
\item
  Input and reference files/directories
\item
  Output files/directories
\item
  Filtering
\item
  Tuning (e.g.~alphas and gammas)
\item
  Seeds
\item
  Any alternatives that you've built-in, e.g.~compress results, use a
  different calculation, verbose output, etc.
\end{itemize}

@@

\fixme{GW}{How to tell if the software is working:}

\begin{itemize}
\item
  Test in a `vanilla' environment, such as another user's computer or a
  dummy account with none of the settings of the original developer.
\item
  Test with different sizes of data: ridiculously small, small, medium,
  and large. The software should run on all sizes given some parameter
  tweaking, or fail with a sensible error message if the data is too
  big.
\item
  Compare results from multiple iterations that have the same parameters
  and inputs.(See rule \#8.)
\item
  Consider a container - but most of these rules still apply
\end{itemize}

>>>>

\begin{flushleft}
{\large\textbf{Ten Simple Rules for Making Research Software More Robust}}
\\
{Morgan~Taschuk}$^{1,\ast}$,
{Greg~Wilson}$^{2}$
\\
1) Ontario Institute for Cancer Research / morgan.taschuk@oicr.on.ca
\\
2) Software Carpentry Foundation / gvwilson@software-carpentry.org
\\
$\ast$ Corresponding author.
\end{flushleft}

\begin{quote}
\begin{center}\textbf{Abstract}\end{center}
Software produced for research,
published and otherwise,
suffers from a number of common problems
that make it difficult or impossible to run outside the original institution,
or even off the primary developer's computer.
We present ten simple rules to make such software robust enough to run anywhere
and thereby delight your users and collaborators.
\end{quote}

@@

Most software has a version number composed of a decimal number that
increments as new versions are released. There are many different ways
to construct and interpret the version, but most importantly for us, a
particular software version run with the same parameters should give
identical results no matter when it's run. Results include both correct
output as well as any errors.
Every time you release your software, i.e., distribute it to
someone other than yourself and/or the development team, you should
increment your version number.

\withurl{Semantic versioning}{http://semver.org/} is one of the most common
types of versioning for open-source software. Version numbers take the
form of \emph{MAJOR.MINOR{[}.PATCH{]}}, e.g., 0.2.6. The major and
minor numbers versions are almost always provided. Changes in the major
version number herald significant changes in the software that are not
backwards compatible, such as changing or removing features or altering
the primary functions of the software. Increasing the minor version
represents incremental improvements in the software, like adding new
features. Following the minor version number can be an arbitrary number
of project-specific identifiers, including patches, builds and qualifiers.
Common qualifiers include \texttt{alpha}, \texttt{beta}, and \texttt{SNAPSHOT},
for applications that are
not yet stable or released, and \texttt{-RC} for release candidates prior
to official release.

@@

The most common way to do this is to include the version number in the
name of the released package, and to link to those old versions from
the project's home page or downloads page.
For example,
as of the time of writing,
the downloads page for the Python programming language includes links to
Versions 3.5.2, 3.5.1, 3.5.0, 3.4.5, 3.4.4, 2.7.12, and 2.7.11,
along with a link to an archive page of even older versions.
Such archives are most useful if these packages document their dependencies,
which we discuss in the next section.

@@

Useful parameters are those values that a user will need to modify to suit
their computer, dataset or application. 
The full list of useful parameters is software-specific and so cannot be
detailed here, but a short list includes
input and reference files and directories,
output files and directories,
filtering,
tuning,
random number generation seeds,
and
any alternatives that you've built-in,
e.g.,
compress results, use a different calculation, verbose output, etc.

@@

There has been extended discussion over the past few years of the
sustainability of research software, but this question is meaningless
in isolation: any piece of software can be sustained if its users are
willing to put in enough effort.  The real equation is the ratio
between the skill and effort available, and the ease with which
software can be installed, understood, used, maintained, and extended.
Following the ten rules we outline here reduce the denominator, and
thereby enable researchers to build on each other's work more easily.
